### 2022-05-06 MSC-GUDMUNDSSON-2022.tex
Added content to methods, added abstract, cleanup, new diagrams and more.

%% ---------------------------------------------------------------
%% $URL: https://repository.cs.ru.is/svn/thesis-template/trunk/ruthesis/latex/DEGREE-NAME-YEAR.tex $
%% $Id: DEGREE-NAME-YEAR.tex 360 2019-02-13 22:04:35Z foley $
%% This is a template LaTeX file for dissertations, theses, or reports at Reykjavík University
%% 
%% Comments and questions can be sent to the RU LaTeX group (latex AT list.ru.is) 
%% ---------------------------------------------------------------

%% METHOD:
%% 0) Read ruthesis/thesis-instructions.pdf
%%    If it is missing, goto https://repository.cs.ru.is/svn/thesis-template/trunk/ruthesis/thesis-instructions.pdf
%% 0.2) Subscribe to the announcements email list at
%%    https://list.ru.is/mailman/listinfo/latex-announcements
%% 1 LaTeX instructions.tex or goto http://afs.rnd.ru.is/project/thesis-template/trunk/ruthesis/latex/instructions.pdf
%% 2) Copy the template files (or unzip) to your working area
%% 3) Rename this file (if needed) with your information e.g. MSC-FOLEY-2007.tex
%% 4) Modify this file to fit your needs (please follow all comments below in the text)
%% 5) For making bibliographies, run "biber".  You can also change
%%    this back to "bibtex".  See below in "Bibliography options".

%%%%%%% CHOOSE ONE OF THESE %%%%%%%%%%%%%%%
%% projectreport: Project report (CS)
%% bachelors: Bachelor of Science thesis
%% masters: Master of Science thesis
%% doctorate: Doctor of Philosophy dissertation
%
%%%%%%% CHOOSE ONE OF THESE %%%%%%
%% 
%% draft: speed up processing by skipping graphics and adding useful
%%     information for editing.  Also sets spacing to double so that it is easier to
%%     write editing marks on paper copy.
%% proof:  proofreading version (final formatting with warnings)
%% final: generate document for submission, removing FIXMEs, and
%%     other markup.  Throw error if any fatal FIXMEs still in document.
%%
%%%%%%% CHOOSE ONE OF THESE IF APPLICABLE %%%%%%
%%
%% deptsse: School of Science and Engineering
%% deptscs: School of Computer Science
%%
%%%%%%%% CHOOSE ANY COMBINATION OF THESE %%%%%%%%%%%%
%%
%% forcegraphics: force graphics, etc. to be included, even in draft mode
%% debug:  writes more messages to the log file, adds debugging output 
%%     and sizing boxes
%% icelandic: thesis is in Icelandic
%% oldstyle:  use the PhD headers and footers from the old CS template
%% online: for online versions (skip blank pages)
\documentclass[online,masters,deptscs,forcegraphics,final]{ruthesis}

%%%%%%%%%%%%%%%%%%%% TeXStudio Magic Comments %%%%%%%%%%%%%%%%%%%%%
%% These comments that start with "!TeX" modify the way TeXStudio works
%% For details see http://texstudio.sourceforge.neit/manual/current/usermanual_en.html   Section 4.10
%%
%% What encoding is the file in?
% !TeX encoding = UTF-8
%% What language should it be spellchecked?
% !TeX spellcheck = en_US
%% What program should I compile this document with?
% !TeX program = xelatex

%%%%%%%%%%%%%%%%%%%% Bibliography options %%%%%%%%%%%%%%%%%%%%%
%% We suggest switching from bibtex to biblatex/biber because it is better able
%% to deal with Icelandic characters and other bibliography issues
%% As long as you use biblatex instead of bibtex by itself, it will at least
%%  generate a document without errors.
%% !!!If you are using TeXStudio, don't forget to update the bibliography setting!!!
\usepackage[backend=biber,bibencoding=utf8,style=ieee]{biblatex}
%\DeclareLanguageMapping{american}{american-apa}  
% need to declare mapping for style=apa to alphabetize properly
% If you set backend=bibtex, it will use bibtex for processing (old way)
%    this can work with Icelandic characters, but you may get weird results.
%    bibtex does not know how to sort Þ and ð
% if you set backend=biber, you can use UTF8 characters such as Þ and
%     ð  but you will have to remember to switch from using bibtex to 
%     biber in your client
% If you use JabRef, make sure the file is encoded in UTF-8 which is
%    not the default.

%% This tells TeXStudio to use biber
% !TeX TXS-program:bibliography = txs:///biber
%% This also sets the bibliography program for TeXShop and TeXWorks
% !BIB program = biber

% Where is your reference library?
\addbibresource{references.bib}
\addbibresource{more_references.bib}

%%%%%%%%%%%%%%%%%%% CUSTOMIZATIONS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% It is not recommended that you customize this file nor
%% ruthesis.cls.  Just fill in the necessary fields.  You should put
%% your macros and packages into a separate file so that it is easier
%% to use updates to the template.  The custom.sty file was created
%% for this reason.  We load this much later so that it can overrite
%% any existing settings
\IfFileExists{custom.sty}{\usepackage{custom}}{}

%%%%%%%%%%%%%%% INFORMATION %%%%%%%%%%%%%%%%%%5
%% University information must be multilingual to deal with the
%%  required cover pages and abstract on thesis
%% NOTE: This may not be required for other reports!!!

%% Babel Icelandic macros are setup  on RedHat at
%% /usr/share/texlive/texmf-dist/tex/generic/babel/icelandic.sty
%% /usr/share/texlive/texmf-dist/tex/generic/babel-icelandic/icelandic.ldf

%% Multilingual macros
%\newML{macroname}{englishword}{icelandicword}
%  creates \macronameML
%    \MLmacroname[english] - returns the english word
%    \MLmacroname[icelandic] - returns the icelandic word
%    \MLmacroname  - uses the current language setting
% Some useful ones have already been defined, but can be redefined
%% Predefined: \MLIceland \MLReykjavikUniversity \MLUniversityIceland

%% What institute?  Default is RU.
%\setInstitution{\MLReykjavikUniversity}
% \newML{InstitutionAddress}{Menntavegur 1\\101 Reykjavík, Iceland}
% {Menntavegi 1\\101 Reykjavík, Ísland}
% \setInstitutionAddress{\MLInstitutionAddress}
% \newML{Tel}{Tel.}{Sími}
% \setInstitutionPhone{\MLTel{} +354 599 6200\\
% Fax +354 599 6201}
% \setInstitutionURL{www.ru.is}

%% ONLY SET DEPARTMENT IF YOU HAVE NOT USED THE deptsse or deptscs OPTION!
%% Department and degree program
%\newML{ND}{New Department}{Nytt deild}
%\setSchool{\MLND}

%% Set your program of study
\newML{program}{Software Engineering}{hugbúnaðarverkfræði}
\program{\MLprogram}

%% Degree long name.  If not already defined, you can create a macro
%\newML{DEGREE}{English Degree Name}{Icelandic Degree Name}
%% Default is set based upon doctorate vs masters option
%% Predefined: \MLMSc \MLPhd
%\setDegreelong{\MLMSc}

%% Degree abb, change if default is not right
%% Default is set based upon doctorate vs masters option
%\degreeabbrv{Sc.D.}

%\setFrontLogo{reyst-logo}
%% Use this if you need a different front logo on the first page
%% e.g. reyst-logo

%% Date in english and icelandic
%% NOTE: THIS IS THE DATE OF THE SUPERVISOR'S SIGNATURE!!!!!!
%% Predefined: \MLjan, \MLfeb, \MLmar, ... \MLdec
%\whensigned{day}{month}{year} %day is only used on some formats, but you must put something.
\whensigned{10}{\MLapr}{2022}

%% Title first in English then Icelandic
%% You need to put both a normal case and ALL CAPS version into the macros.
%%
\newML{Title}{Forecasting Solar Power Production Using Deep Learning and High Resolution Weather Forecasts}{Sólarorkuframleiðsluspár með djúpnámi út frá háupplausnar veðurspám}
\newML{TITLE}{FORECASTING SOLAR POWER PRODUCTION USING DEEP LEARNING AND HIGH RESOLUTION WEATHER FORECASTS}{SÓLARORKUFRAMLEIÐSLUSPÁR MEÐ DJÚPNÁMI ÚT FRÁ HÁUPPLAUSNAR VEÐURSPÁM}
%%
\setTitle{\MLTitle}{\MLTITLE}
%% ***** Special Titles ******
%% If the title must be formatted specifically for the cover page or internal pages
%% (typically via line-breaks using the \newline command) then the following commands must be used 
%%
%\setTitleCover{\MLTITLE}
%% These two for the internal cover pages, usually not needed
%\newML{TitleInternal}{Internal Title}{Icelandic Internal Title}
%\setTitleInternal{\MLTitleInternal}

%% Author name (should be the same in any language, if not use \newML)
%% If you are writing a Project report with multiple authors, separate them with \\:
%% To keep the names typeset together, you want to use non-breaking spaces: ~
%\author{Firstname1~Lastname1\\Firstname2~Lastname2}
\author{Einar~Helgi~Guðmundsson}

%% If the name must be formatted specifically for the signature page
%% (typically via line-breaks) then the following command must be used 
%\setAuthorSignature{Student\\Name}
%% This macro adjusts the author name in the headers of the oldstyle formatting
%\setAuthorHeader{StudentLast}

%%% TODO:  Move the bachelor's form separately -- it confuses people. --foley
%%%%%%%%%%%%%%%%%%%%%%%%%% Project Report or Bachelor's Only!!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setCourse{VT LOK 1012}

%%%%%%%%%%%%%%%%%%%%%%%%%% Bachelors Only!!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setID{160992--2989}%kennitala
\setSemester{2022-2}
\setShortSignedDate{10.4.2022}

\setOrganization{Belgingur ehf.\\Grensásvegur 9\\108 Reykjavík}
\setSubProgram{Hugbúnaðarverkfræði}

%% If the thesis is confidential, uncomment this with the date it can be released
%\setClosedDistribution{10.1.2016}%

%% Put your keywords here in English, then Icelandic.  Separate them with commas.
\newML{keywords}{Deep Learning, Solar Power, Forecast}{Djúpnám, Sólarorka, spá}
\setKeywords{\MLkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%% Masters Only!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% How many credits (ECTS) on Master's degree
%% Usually 30 or 60
\ects{60}

%%%%%%%%%%%%%%%%%%%%%%%%%%% Doctorate Only!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some Computer Science Thesis have an ISSN number.
%% Most other documents do not.
%\bookidnumber{ISSN: 1670-8539} 
%% ID numbers are optional, but nice for sorting in libraries

%% International Standard Book Number (ISBN)
%% This is what most people should use if the thesis is being published.

%% International Standard Serial Number (ISSN)
%% This is usually only for a PhD dissertation as part of a series when published
%%   Computer Science: 1670-8539

%% Additional degrees?  (optional, usually not needed)
%\adddegree{(list of degrees in appendix)}{(sjá lista yfir prófgraður í viðauka)}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% List the entire committee.  Each member has a name (degree should be omitted, unless it is not PhD),
%% Supervisor(s) must appear first
%% On a Bachelors, there is usually only one supervisor and one examiner.

%% Format for each entry:
%%  \personinfo{Name}{Role}{Job Title}{Company/institution}{Country}
%% Predefined macros: \MLSupervisor \MLSupervisors \MLExaminer \MLExaminers

%% Change these to singular/plural as needed.
%% Just uncomment and change the plurality of the macro.
%\setSupervisorHeading{\MLSupervisors}
%\setExaminerHeading{\MLExaminer}

%% Predefined macros:
%% \MLSeniorProfessor \MLProfessor \MLAssociateProfessor \MLAdjunctProfessor \MLEmeritusProfessor \Iceland
%% \MLReykjavikUniversity \MLUniversityIceland

%% Bachelors: primary advisor (Umsjónarkennari), ONLY ONE!
%% All others: As many as you want
\supervisors{
  \personinfo{Dr. Yngvi Björnsson, PhD}{\MLSupervisor}{\MLProfessor}{\MLReykjavikUniversity}{\MLIceland}
  \personinfo{Dr. Ólafur Rögnvaldsson, PhD}{Examiner}{Managing Director}{Belgingur ehf.}{\MLIceland}
%  \personinfo{Ian M. Great}{Co-advisor}{\MLProfessor}{Hochschule Düsseldorf}{Germany}
}

%% Bachelors: secondary advisor (Leiðbeinandi), ONLY ONE
%% All others: As many as you want
\examiners{
  \personinfo{Tough E. Questions}{\MLExaminer}{Associate Professor}{Massachusetts Institute of Technology}{USA}

}

%% An abstract is required to be in both Icelandic and English for most degrees.
%% It is considered good form to limit the abstract to a single paragraph in each language,
%%   at 300 words.  Refer to your degree's instructions.
%% Note: Icelandic quotation marks cannot be typeset using "` and "'.  You should use \enquote{}
%% this is probably due to interactions with the MultiLingual macros.
%% TODO: turn this into more sensible macros to avoid confusion --foley
\newML{AbstractText}{
The world is moving at increasing speed away from generating electricity using fossil fuels, toward more environmentally friendly options, such as solar power. Solar power production is volatile and highly dependent on weather conditions, primarily cloud cover. Instability and unpredictability are undesirable in power production since production must match demand. In order to ensure power availability, highly polluting and expensive fossil power stations, \enquote{peaker plants}, are kept on standby to react to the instability in solar power. The ability to reliably predict solar power production provides more predictability to power grid operators, allowing for greater efficiency in grid operation. 
Valuable data exists in weather forecasts, which can be harnessed to increase predictability in solar power generation. We use novel deep learning methods, attention, and specifically Temporal Fusion Transformers, to forecast the irradiance at a solar power station, which correlates very strongly with power output. We base this on Belgingur ehf.'s high-resolution weather forecast data, the 24-hour forecast irradiance values for the area of the power station.
Our model achieves a very low average Mean Absolute Percentage Error of 13.21\% on the test dataset for one-day forecasts. In addition, the model outputs a confidence level for its predictions providing insight to a grid operator on whether to expect unpredictability in the power output. The model is able to generate more accurate predictions when utilizing forecast data from a wider area around the power station than exclusively data from the power station‘s immediate vicinity.

}  
{
Íslenska!
} % Icelandic abstract goes here
\setAbstract{\MLAbstractText}

%%%%%%%%%%%%%%INDEX SETUP %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Indexes, and other auto-generated material
%% The Memoir package (which we use) automatically generates the index
%% See section 17.2 on page 302 of the guide
%% http://texdoc.net/texmf-dist/doc/latex/memoir/memman.pdf
%% This means you have to run "makeindex DEGREE-NAME-YEAR"
%% !!!Do not load any of the index packages, they cause problems with Memoir!!!
%% !!!You have been warned!!!
%% Note that memoir changes the [] options to only be for filenames, not other options!
\makeindex{}
\indexintoc{}

%% For abbreviations, you may want to try
%% Watch out though, each new index writes another external file and 
%% latex can only write a limited number of them
%%\usepackage[intoc]{nomencl} % intoc: In Table of Contents
%% remember to run:
%% makeindex filename.nlo  -s nomencl.ist -o filename.nls

\finalifforcegraphics{hyperref} %hyperlinks even in draft mode
\usepackage[hidelinks]{hyperref} 
%% !!!Must be the last package loaded except otherwise mentioned!!!!
%% \usepackage{hypcap}  %% puts link at top of figure, must be after hyperref
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{afterpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% DOCUMENT START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%% Some elements have different names on the RU Masters rules
%% They will be annotated with RUM: "name"
\frontmatter{} % setup formatting at beginning

%\frontcover{}%%If you want to see what it looks like with the printed cover
%% TODO:  link to fill-in PDF file on RU website

\frontrequiredpages{}%% the various signaturepages and abstract
%%% WARNING:  if you get an error on the previous line, it is probably because
%%% you put a bad macro or something strange in a title, author, or abstract.

\ifdraft{\coverchapter{Important!!!  Read the Instructions!!!} If you
  have not already done so, \LaTeX{} the \path{instructions.tex} to
  learn how to setup your document and use some of the features.  You
  can see a (somewhat recent) rendered PDF of the instructions included in this folder at \path{instructions-publish.pdf}.
  There is also more information on working with \LaTeX{} at
  \url{http://samvinna.ru.is/project/htgaru/how-to-get-around-projects-publish.pdf}.
  This includes common problems and fixes.

  This page will disappear in anything other than draft mode.}{}

%% Dedication is optional, comment out if it is absent
%% RUM: Not mentioned
\begin{dedications}
  I dedicate this to my spouse/child/pet/power animal.
\end{dedications}

\enableindents{}% turn on/off paragraph indents
% RUM: "Acknowledgements (optional)"
\coverchapter{Acknowledgements} 
\begin{quotation}
So long, and thanks for all the fish.
\end{quotation}\sourceatright{Douglas Adams\cite{adams84fish}}
\vspace{\baselineskip}

\draftnote{Acknowledgements are optional; comment this chapter out if they are absent
  Note that it is important to acknowledge any funding that helped in the work}

The project is partially funded by the Icelandic Technology Development Fund through grant nr. 1910034-0611.

\coverchapter{Preface}
% RUM: "Preface (optional)"
This dissertation is original work by the author, Einar~Helgi~Guðmundsson.

\draftnote{The preface is an optional element
  explaining a little who performed what work.  See
  \url{https://www.grad.ubc.ca/sites/default/files/materials/thesis_sample_prefaces.pdf}
  for suggestions.

  List of publications as part of the preface is
  optional unless elements of the work have already been published.
  It should be a comprehensive list of all publications in which
  material in the thesis has appeared, preferably with references to
  sections as appropriate.  This is also a good place to state
  contribution of student and contribution of others to the work
  represented in the thesis.}

%\coverchapter{Publications}
%% RUM: Not mentioned, this was found in the CS thesis template.  
%% Maybe more applicable to PhD dissertations?
%%% Probably a duplication from before Preface became standard.

\starttables{}% setup formatting
%% TOC, list of figures and list of tables are required
\tableofcontents{}\clearpage%%RUM: "Table of contents"
\listoffigures{}\clearpage%%RUM: "List of figures"
\listoftables{}\clearpage%%RUM: "List of tables"

%\coverchapter{List of drawings and enclosed material}
%RUM: "List of drawings and enclosed material, e.g. CD(as appropriate)"

\listoffixmes{}
% if using fixme package, lists what needs to be done

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
PhD &Doctor of Philosophy\\
ANN &Artificial Neural Network\\
NN &Neural Network\\
RNN &Recurrent Neural Network\\
GRU &Gated Recurrent Unit\\
LSTM &Long Short Term Memory\\
NLP &Natural Language Processing\\
TCNN &Temporal Convolutional Neural Network\\
NARXNN &Non-linear Auto Regressive Neural Network\\
TFT &Temporal Fusion Transformer\\
GRN &Gated Residual Network\\
MAPE &Mean Absolute Percentage Error\\
MAE &Mean Absolute Error\\
CUDA &Compute Unified Device Architecture\\
AMD &Advanced Micro Devices inc.\\
CPU &Central Processing Unit\\
RAM &Random Access Memory\\

\end{tabular}

%\coverchapter{List of Symbols}%%RUM: Not mentioned
%\begin{tabular}{lll}
%Symbol &Description &Value/Units\\
%$E$ &Energy &\si{\joule}\\
%$m$ &Mass &\si{gram}\\
%$c$ &Speed of Light &\SI{2.99E8}{\meter\per\second\square}\\
%\end{tabular}

%% This command prepares for the actual text, e.g. by 
%% calling \mainmatter{}
\starttext{}

%% ---------------------------------------------------------------
%% From this point on, it is standard Latex, except the very end.
%% This is a "report"-based template, so the top-level heading 
%% is \chapter{}

%% WARNING: Make sure that all of these files (and any new ones)
%% are UTF-8 otherwise you will get weird encoding errors.

%% The default division is IMRAD, you may want to divide differently
%% See the introduction for guidance.

\include{introduction}%%RUM: Introduction
\include{methods}%%RUM: "Methods"
\include{results}%%RUM: "Results"
\include{discussion}%%RUM: "Discussion"

%% ---------------------------------------------------------------
\printbibliography{} %%RUM: "References"

%% If appendices are needed, uncomment the following line
%% and include the appendices in separate files
%\appendix{}%%RUM: "Appendicies (as appropriate)
%\input{code} % as an example, perhaps some of your code

%\backmatter{} % Sections after this don't get numbers
%% We prefer that all elements be numbered

%%%%%%%%%%%%% SHOW INDEX %%%%%%%%%%%%%%%%%%
%% Index, optional.  A good idea on longer documents

% You can put instructions at the beginning of the index:
%\renewcommand{\preindexhook}{%
%  The first page number is usually, but not always,
%  the primary reference to the indexed topic.\vskip\onelineskip}

%% You may have to run "makeindex <FILENAME>" to have it be generated
%% Depending upon which package you chose.
%% 
\clearforchapter{}
\printindex{}%%RUM: Not mentioned

%\backcover{}%%RUM: "Back cover (only Phd)
\end{document}

%% ---------------------------------------------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% TeX-engine: xetex
%%% End: